An Extension of LTL with Rules\\ and its Application to Runtime Verification

### 2019-02-05 main.tex

\appendix

@@

\noindent {\bf Sketch of proof of Theorem~\ref{theo2}.}
As we defined the semantics of
EFLTL by translating into QFLTL, QFLTL is at least as
expressive as EFLTL. We show here that 
the expressive power of EFLTL is strictly smaller than that of
QFLTL through the following example. It consists of
a temporal relation\footnote{We abuse notation and
use the same name for the relation and for its
denotation in the logic. Also, $r [ i ]$ has zero
parameters; as noted before, this is a natural extension to the syntax and semantics presented above.} $r$ with only one parameter, so $r [ i ]$
has a Boolean value. We will henceforth say that $r$ occurs
in state $i$. Another relation is $v$, with
two parameters. For each $i$, if $r$ does not appear in the
state, 
then there is a value $a$ from some (infinite) domain,
such that $v ( a, i )$, 
i.e., $v(a)$ appears in $i$.
The property requires that there are no two sets of values appearing within the relation $v$ (as $v(a)$) between successive occurrences of $r$ that are the same.

@@

>>>>

\newpage
\appendix

\section{Appendix}

@@

\noisubsubsection{Sketch of proof of Theorem~\ref{theo2}.}
As we defined the semantics of
EFLTL by translating into QFLTL, QFLTL is at least as
expressive as EFLTL. We show here that 
the expressive power of EFLTL is strictly smaller than that of
QFLTL through the following example. It consists of
a temporal relation\footnote{We abuse notation and
use the same name for the relation and for its
denotation in the logic. Also, $r [ i ]$ has zero
parameters; as noted before, this is a natural extension to the syntax and semantics presented above.} $r$ with only one parameter, so $r [ i ]$
has a Boolean value. We will henceforth say that $r$ occurs
in state $i$. Another relation is $v$, with
two parameters. For each $i$, if $r$ does not appear in the
state, 
then there is a value $a$ from some (infinite) domain,
such that $v ( a, i )$, 
i.e., $v(a)$ appears in $i$.
The property requires that there are no two sets of values appearing within the relation $v$ (as $v(a)$) between successive occurrences of $r$ that are the same.

@@

%%%M
\subsection*{Details of left to right direction of Theorem~\ref{theo1}.}
One can translate any QLTL property into a second order monadic logic in a standard
way~\cite{Thomas}, and then from
second order monadic logic to a deterministic Muller automaton $M = (S, \iota , \Sigma, \Delta, {\cal F})$ (a B\"{u}chi automaton may not always be deterministic), where $S$ is the set of states, $\iota \in S$ is the initial state, $\Sigma$ is the set of inputs, each $\alpha \in \Sigma$ is a minterm (i.e., conjunction of
each variable in the set negated or non-negated) of the variables in $B$, $\Delta : S \times \Sigma \mapsto S$ is the transition function, and ${\cal F} = \{ F_1 , \ldots , F_k \}$ is the accepting set. A muller automaton 
accepts a run if it 
passes infinitely often in exactly
one of the specified accepting sets. To formulate an equivalent ELTL formula,
the auxiliary variables represent the states $S$ of $M$, i.e., $a_s$ represents the state $s \in S$.

%\klaus{The following formulation above is not not clear: ``where an accepted input needs to pass infinitely often in exactly
%one of the specified sets''.}

\noindent
Then we translate $M$ into an ELTL formula
of the form:

\begin{tabbing}
\ \ \ \ \ $\bigvee_{F \in {\cal F}} \bigwedge_{s \in F} ( \Box\Diamond a_s \wedge \bigwedge_{s \not\in F} \Diamond \Box \neg a_s)$
where \\
\ \ \ \ \ \ \ \ \ \ $a_s :=   ( \alpha_{\iota} \wedge \ominus \false ) \vee \bigvee_{\Delta ( s'
, \alpha ) = s } 
(\alpha \wedge \ominus a_{s'}) 
\, :\,  s \in S$
\end{tabbing}

%\klaus{Above: should $s=i \rightarrow \neg \ominus \true$ instead be (note: two changes):
%$\neg \ominus \true \rightarrow a_s=i$
%}
%\klaus{Above: can't judge whether the $\bigwedge_{s \not\in F} \Diamond \Box %\neg a_s)$ is correct.}
\noindent
Basically, the assignments to the auxiliary variables are used to simulate the change of states\footnote{It is possible to use only $\lceil \log ( | S | ) \rceil$
variables, at the expense of longer and more complicated formulas.}, and the main formula expresses the Muller acceptance condition. \qed
